---
title: "Animal Models - Univariate and Bivariate"
author: "Maegwin Bonar"
date: "2026-06-17"
output:
  html_document:
    highlight: breezedark
    css:
      - "styles.css"
    toc: yes
    toc_float: no
    toc_depth: 5
editor_options:
  markdown:
    wrap: 120
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  eval                        = FALSE,
  echo                        = TRUE,
  cache                       = TRUE, # tidy = TRUE,
  class.output                = "bg-success"
)

```

### Introduction

We used the animal model approach to partition the phenotypic variance of migration (VP) into additive genetic variance (VA), permanent environmental effects (VPE) and residual variance (VR) such that VP = VA + VPE + VR conditional on any fixed effects. Table 1 in the manuscript summarizes the fixed and random effects for each animal model.

To test for potential correlation between movement rate and hindleg length we fit a bivariate animal model (de Villemereuil, 2018) with age, year of migration, and log-transformed distance of migration route as fixed effects. Individual identity was included as a random effect in all the migration models to account for permanent environmental effects (de Villemereuil, 2018). 

### Body size traits

```{r}
#03 - run animal models using MCMCglmm
#input - grm.bin files, phenotype files
#output - GMCMCglmm  models

### Packages ----
libs <- c('dplyr','data.table','plyr','OmicKriging', 'gplots', 'MCMCglmm', 'MasterBayes', 'nadiv',
          'igraph', 'Matrix', 'gap', 'mbend', 'matrixcalc', 'AGHmatrix', 'hglm', 'ggplot2',
          'boot','reshape2', 'parallel')
lapply(libs, require, character.only = TRUE)

##### Functions ----
orderDfGrm=function(grm=grm,phen=phen, id=id){
  #make a null dataframe
  d=NULL
  #get the ids of the GRM in the order they appear
  gid=rownames(grm)
  # get the ids from the id column in the phenotype file
  id=as.character(id)
  
  # For loop for creating the new dataframe with the correct number of individuals
  for(i in 1:length(gid)){
    dtmp=c(gid[i],as.numeric(phen$log_weight[id==gid[i]]), as.numeric(phen$log_hindleg[id==gid[i]]), 
           as.numeric(phen$log_chest[id==gid[i]]), as.numeric(phen$log_ifbf[id==gid[i]]), 
           as.numeric(phen$age[id==gid[i]]), as.character(phen$year[id==gid[i]]), 
           as.character(phen$dec_capture[id==gid[i]]))
    d=rbind(d, dtmp)
  }
  
  # save the dataframe
  d=data.frame(d)
  # get the column names right
  names(d)=c("id","log_weight", "log_hindleg", "log_chest", "log_ifbf", "age","year","dec_capture")
  
  # set numeric variables
  d$age <- as.numeric(d$age)
  d$log_weight <- as.numeric(d$log_weight)
  d$log_hindleg <- as.numeric(d$log_hindleg)
  d$log_chest <- as.numeric(d$log_chest)
  d$log_ifbf <- as.numeric(d$log_ifbf)
  
  return(d)
}

##### Read in the data ----
# Phenotypes
phys <- read.csv('input/phys_phenotypes_tf.csv')
# make sure dependent variables are numeric
str(phys)

# GRMs - less than 90% individual missingness
filenames <- list.files("input/gcta_grm/popmap4r08_90_indrm", 
                        pattern="*.grm.bin", full.names = T)

fileid <- substr(filenames,42,nchar(filenames)-8)
# get the base name
grmFileBase <- lapply(filenames, function(x) substr(x,1, nchar(x) - 4))
# use the read_GRMBin function to read in grms
GRM_90 <- lapply(grmFileBase, function(x) read_GRMBin(x))
# make them matrices
GRM_90.matrix <- lapply(GRM_90, as.matrix)
names(GRM_90.matrix) <- fileid
#fix rownames
rownames(GRM_90.matrix[[1]]) <- gsub("0_","",as.character(rownames(GRM_90.matrix[[1]])))
colnames(GRM_90.matrix[[1]]) <- gsub("0_","",as.character(colnames(GRM_90.matrix[[1]])))
rownames(GRM_90.matrix[[2]]) <- gsub("0_","",as.character(rownames(GRM_90.matrix[[2]])))
colnames(GRM_90.matrix[[2]]) <- gsub("0_","",as.character(colnames(GRM_90.matrix[[2]])))
rownames(GRM_90.matrix[[3]]) <- gsub("0_","",as.character(rownames(GRM_90.matrix[[3]])))
colnames(GRM_90.matrix[[3]]) <- gsub("0_","",as.character(colnames(GRM_90.matrix[[3]])))

# Check if it is positive-definite, if not bend it
is.positive.definite(GRM_90.matrix[[1]])

is.positive.definite(GRM_90.matrix[[2]])

is.positive.definite(GRM_90.matrix[[3]])
GRM_90.bend_unf=bend(GRM_90.matrix[[3]])

# Inverse
Ginv_90_nr0.05<-as(solve(GRM_90.matrix[[1]]),"dgCMatrix") 

colnames(Ginv_90_nr0.05)<-rownames(Ginv_90_nr0.05)

Ginv_90_nr0.1<-as(solve(GRM_90.matrix[[2]]),"dgCMatrix") 
colnames(Ginv_90_nr0.1)<-rownames(Ginv_90_nr0.1)

Ginv_90_unf<-as(solve(GRM_90.bend_unf$bent),"dgCMatrix") 
colnames(Ginv_90_unf)<-rownames(Ginv_90_unf)

# list the matrices in the corresponding order to the df
Ginv_90_list <- list()
Ginv_90_list <- list(Ginv_90_nr0.05, Ginv_90_nr0.1, Ginv_90_unf)

#### Making the models -----

# First make sure everything is in the right order and make a new data frame
# Use the oderDfGrm function on the list of matrices

# GRM 90
df_90 <- list()
df_90 <- lapply(GRM_90.matrix, function(x) orderDfGrm(grm = x, phen = phys, id = phys$animal))

# GRM 95
#df_95 <- list()
#df_95 <- lapply(GRM_95.matrix, function(x) orderDfGrm(grm = x, phen = phys, id = phys$animal))

#### Run the models ---

# set priors
prior <-list()
prior <- list(
  R = list(V=1, nu=0.002), 
  G = list(G1 = list(V=1, nu=0.002)))

# log weight
# make an empty list to store the chains
grm90_chains_weight <- list()

for (x in 1:length(df_90)){
  
  # make the list of 4 chains of mcmc models per one grm
  chains_weight <- mclapply(1:4, function(i) {
    MCMCglmm(log_weight ~ age + dec_capture + year, random=~id, family = "gaussian", 
             prior = prior, ginverse = list(id=Ginv_90_list[[x]]), data = df_90[[x]], nitt = 500000, 
             burnin = 10000, thin = 100, start=list(QUASI=FALSE), verbose=FALSE)
  }, mc.cores=4)
  
  # save the chains to the first spot in the list
  grm90_chains_weight[[x]] <- chains_weight
  # name it the name of the df used
  #names(grm90_chains_weight[x]) <- names(df_90[x])
  
  # Tell about progress
  cat('Completed weight model', x, 'of', length(df_90),'\n')

}

# log hindleg
# make an empty list to store the chains
grm90_chains_hleg <- list()

for (x in 1:length(df_90)){

  # make the list of 4 chains of mcmc models per one grm
  chains_hleg <- mclapply(1:4, function(i) {
    MCMCglmm(log_hindleg ~ age + dec_capture + year, random=~id, family = "gaussian", 
             prior = prior, ginverse = list(id=Ginv_90_list[[x]]), data = df_90[[x]], nitt = 500000, 
             burnin = 10000, thin = 100, start=list(QUASI=FALSE), verbose=FALSE)
  }, mc.cores=4)
  
  # save the chains to the first spot in the list
  grm90_chains_hleg[[x]] <- chains_hleg
  # name it the name of the df used
  #names(grm90_chains_hleg[x]) <- names(df_90[x])
  
  # Tell about progress
  cat('Completed hindleg model', x, 'of', length(df_90),'\n')
  
}

# log chest
# make an empty list to store the chains
grm90_chains_chest <- list()

for (x in 1:length(df_90)){

  # make the list of 4 chains of mcmc models per one grm
  chains_chest <- mclapply(1:4, function(i) {
    MCMCglmm(log_chest ~ age + dec_capture+ year, random=~id, family = "gaussian", 
             prior = prior, ginverse = list(id=Ginv_90_list[[x]]), data = df_90[[x]], nitt = 500000, 
             burnin = 10000, thin = 100, start=list(QUASI=FALSE), verbose=FALSE)
  }, mc.cores=4)
  
  # save the chains to the first spot in the list
  grm90_chains_chest[[x]] <- chains_chest
  # name it the name of the df used
  #names(grm90_chains_chest[x]) <- names(df_90[x])
  
  # Tell about progress
  cat('Completed chest model', x, 'of', length(df_90),'\n')
  
}

# log ifbf
# make an empty list to store the chains
grm90_chains_ifbf <- list()

for (x in 1:length(df_90)){

  # make the list of 4 chains of mcmc models per one grm
  chains_ifbf <- mclapply(1:4, function(i) {
    MCMCglmm(log_ifbf ~ age + dec_capture + year, random=~id, family = "gaussian", 
             prior = prior, ginverse = list(id=Ginv_90_list[[x]]), data = df_90[[x]], nitt = 500000, 
             burnin = 10000, thin = 100, start=list(QUASI=FALSE), verbose=FALSE)
  }, mc.cores=4)
  
  # save the chains to the first spot in the list
  grm90_chains_ifbf[[x]] <- chains_ifbf
  # name it the name of the df used
  #names(grm90_chains_ifbf[x]) <- names(df_90[x])
  
  # Tell about progress
  cat('Completed ifbf model', x, 'of', length(df_90),'\n')

}

save.image(file="phys_update.RData")

#END 

```

### Migration traits (univariate models)

```{r}
# 03 - run animal models using MCMCglmm
#input - grm.bin files, phenotype files
#output - GMCMCglmm  models

### Packages ----
libs <- c('dplyr','data.table','plyr','OmicKriging', 'gplots', 'MCMCglmm', 'MasterBayes', 'nadiv',
          'igraph', 'Matrix', 'gap', 'mbend', 'matrixcalc', 'AGHmatrix', 'hglm', 'ggplot2',
          'boot','reshape2', 'parallel')
lapply(libs, require, character.only = TRUE)

##### Functions ----
orderDfGrmMig=function(grm=grm,phen=phen){
  #make a null dataframe
  d=NULL
  #get the ids of the GRM in the order they appear
  gid=data.table(rownames(grm))
  d=merge(x=gid,y=phen, 
               by.x="V1", 
               by.y='animal',
               all.x=TRUE,
               all.y=FALSE)

  # save the dataframe
  d=d[,c("V1", 'filename', "start", "end", "duration", "log_distance", "log_speed","year", 
         'age', 'log_hindleg', 'range'), with=FALSE]
  # get the column names right
  names(d)=c("animal", 'id', "start", "end", "duration", "log_distance", "log_speed","year", 
             'age', 'log_hindleg', 'range')
  
  # set numeric variables
  d$start <- as.numeric(d$start)
  d$end <- as.numeric(d$end)
  d$duration <- as.numeric(d$duration)
  d$log_distance <- as.numeric(d$log_distance)
  d$log_speed <- as.numeric(d$log_speed)
  d$age <- as.numeric(d$age)
  d$log_hindleg <- as.numeric(d$log_hindleg)
  d$range <- as.factor(d$range)
  d$id <- as.character(d$id)
  
  # set character for year
  d$year <- as.factor(d$year)
  
  # make a permanent environmental effects column
  d$pe <- d$id
  
  return(d)
}

##### Read in the data ----
# Phenotypes
mig <- read.csv('input/mig_phenotypes_tf_new_movrate.csv')

# make sure dependent variables are numeric
str(mig)

# GRMs - less than 90% individual missingness
filenames <- list.files("input/gcta_grm/popmap4r08_90_indrm", 
                        pattern="*.grm.bin", full.names = T)

fileid <- substr(filenames,42,nchar(filenames)-8)
# get the base name
grmFileBase <- lapply(filenames, function(x) substr(x,1, nchar(x) - 4))
# use the read_GRMBin function to read in grms
GRM_90 <- lapply(grmFileBase, function(x) read_GRMBin(x))
# make them matrices
GRM_90.matrix <- lapply(GRM_90, as.matrix)
names(GRM_90.matrix) <- fileid
#fix rownames
rownames(GRM_90.matrix[[1]]) <- gsub("0_","",as.character(rownames(GRM_90.matrix[[1]])))
colnames(GRM_90.matrix[[1]]) <- gsub("0_","",as.character(colnames(GRM_90.matrix[[1]])))
rownames(GRM_90.matrix[[2]]) <- gsub("0_","",as.character(rownames(GRM_90.matrix[[2]])))
colnames(GRM_90.matrix[[2]]) <- gsub("0_","",as.character(colnames(GRM_90.matrix[[2]])))
rownames(GRM_90.matrix[[3]]) <- gsub("0_","",as.character(rownames(GRM_90.matrix[[3]])))
colnames(GRM_90.matrix[[3]]) <- gsub("0_","",as.character(colnames(GRM_90.matrix[[3]])))

# Check if it is positive-definite, if not bend it
is.positive.definite(GRM_90.matrix[[1]])

is.positive.definite(GRM_90.matrix[[2]])

is.positive.definite(GRM_90.matrix[[3]])
GRM_90.bend_unf=bend(GRM_90.matrix[[3]])

# Inverse
Ginv_90_nr0.05<-as(solve(GRM_90.matrix[[1]]),"dgCMatrix") 

colnames(Ginv_90_nr0.05)<-rownames(Ginv_90_nr0.05)

Ginv_90_nr0.1<-as(solve(GRM_90.matrix[[2]]),"dgCMatrix") 
colnames(Ginv_90_nr0.1)<-rownames(Ginv_90_nr0.1)

Ginv_90_unf<-as(solve(GRM_90.bend_unf$bent),"dgCMatrix") 
colnames(Ginv_90_unf)<-rownames(Ginv_90_unf)

# list the matrices in the corresponding order to the df
Ginv_90_list <- list()
Ginv_90_list <- list(Ginv_90_nr0.05, Ginv_90_nr0.1, Ginv_90_unf)

#### Making the models -----

# First make sure everything is in the right order and make a new data frame
# Use the oderDfGrm function on the list of matrices

# GRM 90
df_90 <- list()
df_90 <- lapply(GRM_90.matrix, function(x) orderDfGrmMig(grm = x, phen = mig))

#### Run the models ---

# set priors
prior <- list(
  R = list(V=1, nu=0.002), 
  G = list(G1 = list(V=1, nu=0.002), G2 = list(V = 1, nu = 0.002)))

# start(poisson)
# make an empty list to store the chains
grm90_chains_start <- list()

for (x in 1:length(df_90)){
  
  # make the list of 4 chains of mcmc models per one grm
  chains_start <- mclapply(1:4, function(i) {
    MCMCglmm(start ~ range + age + year, random= ~ id + pe, family = "poisson", 
             prior = prior, ginverse = list(animal=Ginv_90_list[[x]]), data = na.omit(df_90[[x]]), nitt = 500000, 
             burnin = 10000, thin = 100, start=list(QUASI=FALSE), verbose=FALSE)
  }, mc.cores=4)
  
  # save the chains to the first spot in the list
  grm90_chains_start[[x]] <- chains_start
  # name it the name of the df used
  #names(grm90_chains_weight[x]) <- names(df_90[x])
  
  # Tell about progress
  cat('Completed start model', x, 'of', length(df_90),'\n')
  
}

# end (poisson)
# make an empty list to store the chains
grm90_chains_end <- list()

for (x in 1:length(df_90)){
  
  # make the list of 4 chains of mcmc models per one grm
  chains_end <- mclapply(1:4, function(i) {
    MCMCglmm(end ~ range + age + year, random=~ id + pe, family = "poisson", 
             prior = prior, ginverse = list(animal=Ginv_90_list[[x]]), data = na.omit(df_90[[x]]), nitt = 500000, 
             burnin = 10000, thin = 100, start=list(QUASI=FALSE), verbose=FALSE)
  }, mc.cores=4)
  
  # save the chains to the first spot in the list
  grm90_chains_end[[x]] <- chains_end
  # name it the name of the df used
  #names(grm90_chains_weight[x]) <- names(df_90[x])
  
  # Tell about progress
  cat('Completed end model', x, 'of', length(df_90),'\n')
  
}

# duration (poisson)
# make an empty list to store the chains
grm90_chains_duration <- list()

for (x in 1:length(df_90)){
  
  # make the list of 4 chains of mcmc models per one grm
  chains_duration <- mclapply(1:4, function(i) {
    MCMCglmm(duration ~ range + age + log_distance + year, random=~ id + pe, family = "poisson", 
             prior = prior, ginverse = list(animal=Ginv_90_list[[x]]), data = na.omit(df_90[[x]]), nitt = 500000, 
             burnin = 10000, thin = 100, start=list(QUASI=FALSE), verbose=FALSE)
  }, mc.cores=4)
  
  # save the chains to the first spot in the list
  grm90_chains_duration[[x]] <- chains_duration
  # name it the name of the df used
  #names(grm90_chains_weight[x]) <- names(df_90[x])
  
  # Tell about progress
  cat('Completed duration model', x, 'of', length(df_90),'\n')
  
}

# log_speed MIGRATORY
# make an empty list to store the chains
grm90_chains_speed_dist <- list()

for (x in 1:length(df_90)){
  
  # make the list of 4 chains of mcmc models per one grm
  chains_speed <- mclapply(1:4, function(i) {
    MCMCglmm(log_speed2 ~ log_distance + age + year, random=~ id + pe, family = "gaussian", 
             prior = prior, ginverse = list(animal=Ginv_90_list[[x]]), data = na.omit(df_90[[x]]), nitt = 500000, 
             burnin = 10000, thin = 100, start=list(QUASI=FALSE), verbose=FALSE)
  }, mc.cores=4)
  
  # save the chains to the first spot in the list
  grm90_chains_speed_dist[[x]] <- chains_speed
  # name it the name of the df used
  #names(grm90_chains_weight[x]) <- names(df_90[x])
  
  # Tell about progress
  cat('Completed speed model', x, 'of', length(df_90),'\n')
  
}

# log_speed MARCH NON-MIGRATORY
# make an empty list to store the chains
grm90_chains_mar_mov <- list()

for (x in 1:length(df_90)){
  
  # make the list of 4 chains of mcmc models per one grm
  chains_mar_mov <- mclapply(1:4, function(i) {
    MCMCglmm(log_mar_speed ~ age + year, random=~ id + pe, family = "gaussian", 
             prior = prior, ginverse = list(animal=Ginv_90_list[[x]]), data = na.omit(df_90[[x]]), nitt = 500000, 
             burnin = 10000, thin = 100, start=list(QUASI=FALSE), verbose=FALSE)
  }, mc.cores=4)
  
  # save the chains to the first spot in the list
  grm90_chains_mar_mov[[x]] <- chains_mar_mov
  # name it the name of the df used
  #names(grm90_chains_weight[x]) <- names(df_90[x])
  
  # Tell about progress
  cat('Completed March speed model', x, 'of', length(df_90),'\n')
  
}

# log_distance
# make an empty list to store the chains
grm90_chains_dist <- list()

for (x in 1:length(df_90)){
  
  # make the list of 4 chains of mcmc models per one grm
  chains_dist <- mclapply(1:4, function(i) {
    MCMCglmm(log_distance ~ range + age + year, random= ~ id + pe, family = "gaussian", 
             prior = prior, ginverse = list(animal=Ginv_90_list[[x]]), data = na.omit(df_90[[x]]), nitt = 500000, 
             burnin = 10000, thin = 100, start=list(QUASI=FALSE), verbose=FALSE)
  }, mc.cores=4)
  
  # save the chains to the first spot in the list
  grm90_chains_dist[[x]] <- chains_dist
  # name it the name of the df used
  #names(grm90_chains_weight[x]) <- names(df_90[x])
  
  # Tell about progress
  cat('Completed distance model', x, 'of', length(df_90),'\n')
  
}

save.image(file="mig_update.RData")

#END 
```

### Migration traits (bivariate model)

```{r}
# 03 - run bivariate animal models using MCMCglmm
#input - grm.bin files, phenotype files
#output - GMCMCglmm  bivariate models
# following https://diogro.github.io/QGcourse/tutorials/bivariate_animal_model.html

## NOTE: This is run for all 3 of the GRMs filtered based on relatedness: r=0.05, r=0.01, unfiltered
## r=0.1 is the one we present in the paper

### Packages ----
libs <- c('dplyr','data.table','plyr','OmicKriging', 'gplots', 'MCMCglmm', 'MasterBayes', 'nadiv',
          'igraph', 'Matrix', 'gap', 'mbend', 'matrixcalc', 'AGHmatrix', 'hglm', 'ggplot2',
          'boot','reshape2', 'parallel')
lapply(libs, require, character.only = TRUE)

##### Functions ----
orderDfGrmMig=function(grm=grm,phen=phen){
  #make a null dataframe
  d=NULL
  #get the ids of the GRM in the order they appear
  gid=data.table(rownames(grm))
  d=merge(x=gid,y=phen, 
          by.x="V1", 
          by.y='animal',
          all.x=TRUE,
          all.y=FALSE)
  
  # save the dataframe
  d=d[,c("V1", 'filename', "start", "end", "duration", "log_distance", "log_speed2","log_mar_speed","year", 
         'age', 'log_hindleg', 'range'), with=FALSE]
  # get the column names right
  names(d)=c("animal", 'id', "start", "end", "duration", "log_distance", "log_speed2","log_mar_speed","year", 
             'age', 'log_hindleg', 'range')
  
  # set numeric variables
  d$start <- as.numeric(d$start)
  d$end <- as.numeric(d$end)
  d$duration <- as.numeric(d$duration)
  d$log_distance <- as.numeric(d$log_distance)
  d$log_speed <- as.numeric(d$log_speed2)
  d$log_mar_speed <- as.numeric(d$log_mar_speed)
  d$age <- as.numeric(d$age)
  d$log_hindleg <- as.numeric(d$log_hindleg)
  d$range <- as.factor(d$range)
  d$id <- as.character(d$id)
  d$animalid <- as.character(d$id)
  
  # make a permanent environmental effects column
  d$pe <- d$id
  
  # set character for year
  d$year <- as.factor(d$year)
  
  return(d)
}


##### Read in the data ----
# Phenotypes
mig <- read.csv('input/mig_phenotypes_tf_new_movrate.csv')

# make sure dependent variables are numeric
str(mig)

# GRMs - less than 90% individual missingness
filenames <- list.files("input/gcta_grm/popmap4r08_90_indrm", 
                        pattern="*.grm.bin", full.names = T)

fileid <- substr(filenames,42,nchar(filenames)-8)
# get the base name
grmFileBase <- lapply(filenames, function(x) substr(x,1, nchar(x) - 4))
# use the read_GRMBin function to read in grms
GRM_90 <- lapply(grmFileBase, function(x) read_GRMBin(x))
# make them matrices
GRM_90.matrix <- lapply(GRM_90, as.matrix)
names(GRM_90.matrix) <- fileid
#fix rownames
rownames(GRM_90.matrix[[1]]) <- gsub("0_","",as.character(rownames(GRM_90.matrix[[1]])))
colnames(GRM_90.matrix[[1]]) <- gsub("0_","",as.character(colnames(GRM_90.matrix[[1]])))
rownames(GRM_90.matrix[[2]]) <- gsub("0_","",as.character(rownames(GRM_90.matrix[[2]])))
colnames(GRM_90.matrix[[2]]) <- gsub("0_","",as.character(colnames(GRM_90.matrix[[2]])))
rownames(GRM_90.matrix[[3]]) <- gsub("0_","",as.character(rownames(GRM_90.matrix[[3]])))
colnames(GRM_90.matrix[[3]]) <- gsub("0_","",as.character(colnames(GRM_90.matrix[[3]])))

# Check if it is positive-definite, if not bend it
is.positive.definite(GRM_90.matrix[[1]])

is.positive.definite(GRM_90.matrix[[2]])

is.positive.definite(GRM_90.matrix[[3]])
GRM_90.bend_unf=bend(GRM_90.matrix[[3]])

# Inverse
Ginv_90_nr0.05<-as(solve(GRM_90.matrix[[1]]),"dgCMatrix") 

colnames(Ginv_90_nr0.05)<-rownames(Ginv_90_nr0.05)

Ginv_90_nr0.1<-as(solve(GRM_90.matrix[[2]]),"dgCMatrix") 
colnames(Ginv_90_nr0.1)<-rownames(Ginv_90_nr0.1)

Ginv_90_unf<-as(solve(GRM_90.bend_unf$bent),"dgCMatrix") 
colnames(Ginv_90_unf)<-rownames(Ginv_90_unf)

# list the matrices in the corresponding order to the df
Ginv_90_list <- list()
Ginv_90_list <- list(Ginv_90_nr0.05, Ginv_90_nr0.1, Ginv_90_unf)

#### Making the models -----

#1. Test for speed-hindleg relationship
mod_regression<-lm(mig$log_speed2~mig$log_hindleg) # data=voles)
plot(mig$log_speed2,mig$log_hindleg)               #quick plot in base graphics
abline(mod_regression)

summary(mod_regression)

#2. Fitting a bivariate model

# First make sure everything is in the right order and make a new data frame
# Use the oderDfGrm function on the list of matrices

# GRM 90
df_90 <- list()
df_90 <- lapply(GRM_90.matrix, function(x) orderDfGrmMig(grm = x, phen = mig))

# set priors
prior_bi <- list(R = list(V=diag(2), nu=0.002),
                 G = list(G1 = list(V=diag(2), nu=0.002), G2 = list(V =diag(2), nu = 0.002)))

# make an empty list to store the chains
grm90_chains_speed_bi <- list()

for (x in 1:length(df_90)){
  
  # make the list of 4 chains of mcmc models per one grm
  chains_speed <- mclapply(1:4, function(i) {
    MCMCglmm(cbind(log_speed2, log_hindleg) ~ trait + trait:age +trait:year +trait:log_distance-1, 
             random = ~us(trait):id + us(trait):pe,
             rcov = ~us(trait):units, family = c("gaussian", "gaussian"),
             ginverse = list(animal=Ginv_90_list[[x]]),
             data = na.omit(df_90[[x]]), 
             prior = prior_bi, 
             nitt = 500000, burnin = 10000, thin = 100, 
             start=list(QUASI=FALSE), 
             verbose=FALSE)
  }, mc.cores=4)
  
  # save the chains to the first spot in the list
  grm90_chains_speed_bi[[x]] <- chains_speed
  # name it the name of the df used
  #names(grm90_chains_weight[x]) <- names(df_90[x])
  
  # Tell about progress
  cat('Completed speed:hindleg model', x, 'of', length(df_90),'\n')
  
}

save.image(file="mig_bivariate.RData")

```

