---
title: "Calculating heritability"
author: "Maegwin Bonar"
date: "2026-06-17"
output:
  html_document:
    highlight: breezedark
    css:
      - "styles.css"
    toc: yes
    toc_float: no
    toc_depth: 5
editor_options:
  markdown:
    wrap: 120
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  eval                        = FALSE,
  echo                        = TRUE,
  cache                       = TRUE, # tidy = TRUE,
  class.output                = "bg-success"
)

```

### Introduction

Narrow sense heritability conditional on fixed effects was estimated as h2 = VA/(VA + VPE + VR + Vfixed) and the permanent environmental effect conditional on fixed effects as pe2 = VPE/(VA + VPE + VR + Vfixed). We measured individual repeatability as the ratio of among-individual variance (genetic and nongenetic) over the total phenotypic variance ind2 = VA+VPE/ (VA + VPE + VR+ Vfixed) (Nakagawa & Schielzeth, 2010). Finally, we estimated evolvability for each migration trait (IA) (Hansen et al., 2011; Houle, 1992). Evolvability defined as the mean standardized proportion of additive genetic variance, was calculated as IA = VA/m2, where m is the trait mean (Hansen et al., 2011). 


### Load workspace and functions

```{r}
# 04 - get model output and heritability estimates from mcmcglmm chains
#input - mcmcglmm.RData workspace saved from step 03
#output - heritabiltiy estimates and model ouputs


### Packages ----
libs <- c('dplyr','data.table','plyr','OmicKriging', 'gplots', 'MCMCglmm', 'MasterBayes', 'nadiv',
          'igraph', 'Matrix', 'gap', 'mbend', 'matrixcalc', 'AGHmatrix', 'hglm', 'ggplot2',
          'boot','reshape2', 'parallel', 'readxl', 'QGglmm', 'matrixStats', 'tibble')

lapply(libs, require, character.only = TRUE)

#### Functions ----
ReadGRMBin=function(prefix, AllN=F, size=4){
  sum_i=function(i){
    return((1+i)*i/2)
  }
  
  BinFileName=paste(prefix,".grm.bin",sep="")
  NFileName=paste(prefix,".grm.N.bin",sep="")
  IDFileName=paste(prefix,".grm.id",sep="")
  id = read.table(IDFileName)
  n=dim(id)[1]
  BinFile=file(BinFileName, "rb");
  grm=readBin(BinFile, n=n*(n+1)/2, what=numeric(0), size=size)
  NFile=file(NFileName, "rb");
  if(AllN==T){
    N=readBin(NFile, n=n*(n+1)/2, what=numeric(0), size=size)
  }
  else N=readBin(NFile, n=1, what=numeric(0), size=size)
  i=sapply(1:n, sum_i)
  return(list(diag=grm[i], off=grm[-i], id=id, N=N))
}


##### load workspace ----
load("~/Dropbox/Shafer Lab/RADseq/mig_update.RData")
load("~/Dropbox/Shafer Lab/RADseq/phys_update.RData")

##### Model outputs

summaryOfModels=function(grm, grm.name, name){
# Fixed effects
fixed <- lapply(grm, function(m) m$Sol) # Sol = fixed effects
fixed <- do.call(mcmc.list, fixed)
fixed_all <- do.call(rbind, fixed)

mean_fixed <- colMeans(fixed_all)
quantiles_fixed <- t(colQuantiles(fixed_all, probs = c(0.025, 0.975)))

summary_fixed <- cbind(
  estimate = mean_fixed,
  t(quantiles_fixed),
  trait = paste0(grm.name, "_", name)
)

#a <- summary(fixed)

# Random effects
random <- lapply(grm, function(m) m$VCV) # VCV = variance
random <- do.call(mcmc.list, random)

b <- summary(random)

# Gelman Rubed diagnostics
c <- gelman.diag(random)
d <- gelman.diag(fixed)

# Autocorrelation of diagonals
e <- autocorr.diag(random)
f <- autocorr.diag(fixed)

#return(list(fixed, random, mean_fixed, quantiles_fixed,  b, c, d, e, f))
return(list(summary_fixed, b, c, d, e, f))

}

propVarRandom=function(grm, grm.name, name){
  
  # Random effects
  random <- lapply(grm, function(m) m$VCV) # VCV = variance
  random_mcmc_list <- do.call(mcmc.list, random)
  
  sol <- lapply(grm, function(m) m$Sol)
  sol_all <- do.call(rbind, sol)
  
  # # variance due to fixed effects
  pred <- as.matrix(grm[[1]]$X) %*% t(sol_all)  # predictions from fixed effects
  V_fixed <- apply(pred, 2, var)  
  
  # rbind all chains into one large posterior sample
  vcv_all <- do.call(rbind, random_mcmc_list)
  
  VA  <- vcv_all[,"id"]
  VPE <- vcv_all[,"pe"]
  VR  <- vcv_all[,"units"]
  
  # total
  V_total_cond <- VA + VPE + VR + V_fixed
  # proportions of variance
  prop_VA  <- VA  / V_total_cond
  prop_VPE <- VPE / V_total_cond
  prop_VR  <- VR  / V_total_cond
  prop_fixed <- V_fixed / V_total_cond

  #summary
  summary_df <- data.frame(
    grm = c(paste0(grm.name, "_", name)),
    effect = c("id", "pe", "units", "fixed"),
    mean   = c(mean(prop_VA), mean(prop_VPE), mean(prop_VR), mean(prop_fixed)),
    median = c(median(prop_VA), median(prop_VPE), median(prop_VR), median(prop_fixed)),
    lower  = c(quantile(prop_VA, 0.025),
               quantile(prop_VPE, 0.025),
               quantile(prop_VR, 0.025),
               quantile(prop_fixed, 0.025)),
    upper  = c(quantile(prop_VA, 0.975),
               quantile(prop_VPE, 0.975),
               quantile(prop_VR, 0.975),
               quantile(prop_fixed, 0.975))
  )
  return(summary_df)
}

propVarNoPE=function(grm, grm.name, name){
  
  # Random effects
  random <- lapply(grm, function(m) m$VCV) # VCV = variance
  random_mcmc_list <- do.call(mcmc.list, random)
  
  sol <- lapply(grm, function(m) m$Sol)
  sol_all <- do.call(rbind, sol)
  
  # # variance due to fixed effects
  pred <- as.matrix(grm[[1]]$X) %*% t(sol_all)  # predictions from fixed effects
  V_fixed <- apply(pred, 2, var)  
  
  # rbind all chains into one large posterior sample
  vcv_all <- do.call(rbind, random_mcmc_list)
  
  VA  <- vcv_all[,"id"]
  VR  <- vcv_all[,"units"]
  
  # total
  V_total_cond <- VA + VR + V_fixed
  # proportions of variance
  prop_VA  <- VA  / V_total_cond
  prop_VR  <- VR  / V_total_cond
  prop_fixed <- V_fixed / V_total_cond
  
  #summary
  summary_df <- data.frame(
    grm = c(paste0(grm.name, "_", name)),
    effect = c("id", "units", "fixed"),
    mean   = c(mean(prop_VA),  mean(prop_VR), mean(prop_fixed)),
    median = c(median(prop_VA), median(prop_VR), median(prop_fixed)),
    lower  = c(quantile(prop_VA, 0.025),
               quantile(prop_VR, 0.025),
               quantile(prop_fixed, 0.025)),
    upper  = c(quantile(prop_VA, 0.975),
               quantile(prop_VR, 0.975),
               quantile(prop_fixed, 0.975))
  )
  return(summary_df)
}

propVarRaw=function(grm, grm.name, name){
  
  # Random effects
  random <- lapply(grm, function(m) m$VCV) # VCV = variance
  random_mcmc_list <- do.call(mcmc.list, random)
  
  sol <- lapply(grm, function(m) m$Sol)
  sol_all <- do.call(rbind, sol)
  
  # # variance due to fixed effects
  pred <- as.matrix(grm[[1]]$X) %*% t(sol_all)  # predictions from fixed effects
  V_fixed <- apply(pred, 2, var)  
  
  # rbind all chains into one large posterior sample
  vcv_all <- do.call(rbind, random_mcmc_list)
  
  VA  <- vcv_all[,"id"]
  VPE <- vcv_all[,"pe"]
  VR  <- vcv_all[,"units"]
  
  # total
  V_total_cond <- VA + VPE + VR + V_fixed
  # proportions of variance
  prop_VA  <- VA  / V_total_cond
  prop_VPE <- VPE / V_total_cond
  prop_VR  <- VR  / V_total_cond
  prop_fixed <- V_fixed / V_total_cond
  
  return(list(prop_VA, prop_VPE, prop_VR, prop_fixed))
}

```


### Summary of models and proportion of variance for body size traits

```{r}
# Weight
summary_weight <- list()
for (i in 1:length(grm90_chains_weight)){
  summary_weight[[i]] <- summaryOfModels(grm = grm90_chains_weight[[i]], 
                                         grm.name = 'grm90_chains_weight',
                                         name = i)
}

summary_weight_df <- summary_weight %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

# Hindleg
summary_hleg <- list()
for (i in 1:length(grm90_chains_hleg)){
  summary_hleg[[i]] <- summaryOfModels(grm = grm90_chains_hleg[[i]], 
                                         grm.name = 'grm90_chains_hleg',
                                         name = i)
}

summary_hleg_df <- summary_hleg %>% 
  lapply(as.data.frame) %>% 
  bind_rows()


# Chest
summary_chest <- list()
for (i in 1:length(grm90_chains_chest)){
  summary_chest[[i]] <- summaryOfModels(grm = grm90_chains_chest[[i]], 
                                         grm.name = 'grm90_chains_chest',
                                         name = i)
}

summary_chest_df <- summary_chest %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

# IFBF
summary_ifbf <- list()
for (i in 1:length(grm90_chains_ifbf)){
  summary_ifbf[[i]] <- summaryOfModels(grm = grm90_chains_ifbf[[i]], 
                                         grm.name = 'grm90_chains_ifbf',
                                         name = i)
}

summary_ifbf_df <- summary_ifbf %>% 
  lapply(as.data.frame) %>% 
  bind_rows()


summary_body_all <- rbind(summary_weight_df,summary_chest_df, summary_hleg_df, summary_ifbf_df)

write.csv(summary_body_all, "output/summary_fixed_body_all.csv")

#### Proportion of Variance ---------

# chest
propvar_chest<-list()
for (i in 1:length(grm90_chains_chest)){
  propvar_chest[[i]] <- propVarNoPE(grm = grm90_chains_chest[[i]], 
                                      grm.name = 'grm90_chains_chest',
                                      name = i)
}
propvar_chest_df <- propvar_chest %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

# hindleg
propvar_hleg<-list()
for (i in 1:length(grm90_chains_hleg)){
  propvar_hleg[[i]] <- propVarNoPE(grm = grm90_chains_hleg[[i]], 
                                    grm.name = 'grm90_chains_hleg',
                                    name = i)
}
propvar_hleg_df <- propvar_hleg %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

# ifbf
propvar_ifbf<-list()
for (i in 1:length(grm90_chains_ifbf)){
  propvar_ifbf[[i]] <- propVarNoPE(grm = grm90_chains_ifbf[[i]], 
                                    grm.name = 'grm90_chains_ifbf',
                                    name = i)
}
propvar_ifbf_df <- propvar_ifbf %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

# weight
propvar_weight<-list()
for (i in 1:length(grm90_chains_weight)){
  propvar_weight[[i]] <- propVarNoPE(grm = grm90_chains_weight[[i]], 
                                    grm.name = 'grm90_chains_weight',
                                    name = i)
}
propvar_weight_df <- propvar_weight %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

propvar_body <- rbind(propvar_chest_df, propvar_hleg_df, propvar_ifbf_df, propvar_weight_df)

write.csv(propvar_body, "output/propvar_body.csv")


```


### Summary of models and proportion of variance for migration traits

```{r}

# start
summary_start<- list()
for (i in 1:length(grm90_chains_start)){
  summary_start[[i]] <- summaryOfModels(grm = grm90_chains_start[[i]], 
                                       grm.name = 'grm90_chains_start',
                                       name = i)
}
summary_start_df <- summary_start %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

# proportion of variance for start
propvar_start<-list()
for (i in 1:length(grm90_chains_start)){
  propvar_start[[i]] <- propVarRandom(grm = grm90_chains_start[[i]], 
                                        grm.name = 'grm90_chains_start',
                                        name = i)
}
propvar_start_df <- propvar_start %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

# end
summary_end<- list()
for (i in 1:length(grm90_chains_end)){
  summary_end[[i]] <- summaryOfModels(grm = grm90_chains_end[[i]], 
                                        grm.name = 'grm90_chains_end',
                                        name = i)
}

summary_end_df <- summary_end %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

# proportion of variance for end
propvar_end<-list()
for (i in 1:length(grm90_chains_end)){
  propvar_end[[i]] <- propVarRandom(grm = grm90_chains_end[[i]], 
                                      grm.name = 'grm90_chains_end',
                                      name = i)
}
propvar_end_df <- propvar_end %>% 
  lapply(as.data.frame) %>% 
  bind_rows()


# speed (migratory)
summary_speed<- list()
for (i in 1:length(grm90_chains_speed_dist)){
  summary_speed[[i]] <- summaryOfModels(grm = grm90_chains_speed_dist[[i]], 
                                      grm.name = 'grm90_chains_speed_dist',
                                      name = i)
}

summary_speed_df <- summary_speed %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

# proportion of variance for speed (migratory)
propvar_speed<-list()
for (i in 1:length(grm90_chains_speed_dist)){
  propvar_speed[[i]] <- propVarRandom(grm = grm90_chains_speed_dist[[i]], 
                                      grm.name = 'grm90_chains_speed_dist',
                                      name = i)
}
propvar_speed_df <- propvar_speed %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

# duration
summary_duration<- list()
for (i in 1:length(grm90_chains_duration)){
  summary_duration[[i]] <- summaryOfModels(grm = grm90_chains_duration[[i]], 
                                      grm.name = 'grm90_chains_duration',
                                      name = i)
}

summary_duration_df <- summary_duration %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

# proportion of variance for duration
propvar_duration<-list()
for (i in 1:length(grm90_chains_duration)){
  propvar_duration[[i]] <- propVarRandom(grm = grm90_chains_duration[[i]], 
                                      grm.name = 'grm90_chains_duration',
                                      name = i)
}
propvar_duration_df <- propvar_duration %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

# distance
summary_dist<- list()
for (i in 1:length(grm90_chains_dist)){
  summary_dist[[i]] <- summaryOfModels(grm = grm90_chains_dist[[i]], 
                                      grm.name = 'grm90_chains_dist',
                                      name = i)
}

summary_dist_df <- summary_dist %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

# proportion of variance for distance
propvar_dist<-list()
for (i in 1:length(grm90_chains_dist)){
  propvar_dist[[i]] <- propVarRandom(grm = grm90_chains_dist[[i]], 
                                      grm.name = 'grm90_chains_dist',
                                      name = i)
}
propvar_dist_df <- propvar_dist %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

# March movement rate (non-migratory)
summary_mar_movrate<- list()
for (i in 1:length(grm90_chains_mar_mov)){
  summary_mar_movrate[[i]] <- summaryOfModels(grm = grm90_chains_mar_mov[[i]], 
                                       grm.name = 'grm90_chains_mar_mov',
                                       name = i)
}

summary_marmov_df <- summary_mar_movrate %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

# proportion of variance for non-migratory movement rate
propvar_mar_mov<-list()
for (i in 1:length(grm90_chains_mar_mov)){
  propvar_mar_mov[[i]] <- propVarRandom(grm = grm90_chains_mar_mov[[i]], 
                                      grm.name = 'grm90_chains_mar_mov',
                                      name = i)
}
propvar_mar_mov_df <- propvar_mar_mov %>% 
  lapply(as.data.frame) %>% 
  bind_rows()


summary_fixed_mig <- rbind(summary_start_df, summary_end_df, summary_dist_df, summary_duration_df,
                           summary_speed_df, summary_marmov_df)

write.csv(summary_fixed_mig, "output/summary_fixed_mig_all.csv")

propvar_mig <- rbind(propvar_start_df, propvar_end_df, propvar_dist_df, propvar_duration_df,
                     propvar_speed_df, propvar_mar_mov_df)

write.csv(propvar_mig, "output/propvar_mig.csv")

```

### Heritability for body size traits

```{r}
getHeritabilityPhys = function(grm, grm.name, name){
  
  # Random effects
  random <- lapply(grm, function(m) m$VCV) # VCV = variance
  random <- do.call(mcmc.list, random)
  
  # variance due to fixed effects
  predictions <- lapply(grm, function(m) as.matrix(m$X) %*% t(m$Sol))
  fixedf_variance <- lapply(predictions, function (j) apply(j, MARGIN = 2, var))
  
  # make a dataframe of all the id and untis from all 4 chains
  random_out<-do.call(rbind.data.frame, random)
  fixed_eff <- data.frame(values = as.numeric(unlist(fixedf_variance)))
  
  ## Pull out variance components
  h2 <- random_out$id /
    (random_out$id + random_out$units + fixed_eff$values)
  
  # ind2 <- (random_out$year + random_out$id) /
  #   (random_out$id + random_out$year + random_out$units)
  
  # pe2 <- random_out$year /
  #   (random_out$id + random_out$year + random_out$units)
  
  h2_df <- data.frame(var = 'h2',
                      median = median(h2),
                      mean = mean(h2),
                      lower = quantile(h2, 0.025),
                      upper = quantile(h2, 0.975),
                      grm = paste(grm.name, name, sep = "_"),
                      stringsAsFactors = FALSE)
  
  # pe2_df <- data.frame(var = 'pe2',
  #                     median = median(pe2),
  #                     mean = mean(pe2),
  #                     lower = quantile(pe2, 0.025),
  #                     upper = quantile(pe2, 0.975),
  #                     grm = paste(grm.name, name, sep = "_"),
  #                     stringsAsFactors = FALSE)
  # 
  # ind2_df <- data.frame(var = 'ind2',
  #                     median = median(ind2),
  #                     mean = mean(ind2),
  #                     lower = quantile(ind2, 0.025),
  #                     upper = quantile(ind2, 0.975),
  #                     grm = paste(grm.name, name, sep = "_"),
  #                     stringsAsFactors = FALSE)
  
  #return(list(h2_df, pe2_df, ind2_df))
  return(list(h2_df))
  
}

# weight
h2_weight <- list()
for (i in 1:length(grm90_chains_weight)){
  h2_weight[[i]] <- getHeritabilityPhys(grm = grm90_chains_weight[[i]], 
                                         grm.name = 'grm90_chains_weight',
                                         name = i)
}

nr0.05_weight <- do.call('rbind', h2_weight[[1]])
nr0.1_weight <- do.call('rbind', h2_weight[[2]])
unf_weight <- do.call('rbind', h2_weight[[3]])
df_weight <- h2_weight %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

# hindleg
h2_hleg <- list()
for (i in 1:length(grm90_chains_hleg)){
  h2_hleg[[i]] <- getHeritabilityPhys(grm = grm90_chains_hleg[[i]], 
                                    grm.name = 'grm90_chains_hleg',
                                    name = i)
}

nr0.05_hleg <- do.call('rbind', h2_hleg[[1]])
nr0.1_hleg <- do.call('rbind', h2_hleg[[2]])
unf_hleg <- do.call('rbind', h2_hleg[[3]])
df_hleg <- h2_hleg %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

# chest
h2_chest <- list()
for (i in 1:length(grm90_chains_chest)){
  h2_chest[[i]] <- getHeritabilityPhys(grm = grm90_chains_chest[[i]], 
                                    grm.name = 'grm90_chains_chest',
                                    name = i)
}

nr0.05_chest <- do.call('rbind', h2_chest[[1]])
nr0.1_chest <- do.call('rbind', h2_chest[[2]])
unf_chest <- do.call('rbind', h2_chest[[3]])
df_chest <- h2_chest %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

# IFBF
h2_ifbf <- list()
for (i in 1:length(grm90_chains_ifbf)){
  h2_ifbf[[i]] <- getHeritabilityPhys(grm = grm90_chains_ifbf[[i]], 
                                    grm.name = 'grm90_chains_ifbf',
                                    name = i)
}

nr0.05_ifbf <- do.call('rbind', h2_ifbf[[1]])
nr0.1_ifbf <- do.call('rbind', h2_ifbf[[2]])
unf_ifbf <- do.call('rbind', h2_ifbf[[3]])
df_ifbf <- h2_ifbf %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

df_phys <- rbind(df_weight, df_chest, df_hleg, df_ifbf)
write.csv(df_phys, "output/h2_phys_update.csv")

```

### Heritability, repeatability, evolvability for migration traits

```{r}

getHeritabilityMig = function(grm, grm.name, name, trait){
  
  # Random effects
  random <- lapply(grm, function(m) m$VCV) # VCV = variance
  random <- do.call(mcmc.list, random)
  
  # make a dataframe of all the id and untis from all 4 chains
  random_out<-do.call(rbind.data.frame, random)
  
  # # variance due to fixed effects
  predictions <- lapply(grm, function(m) as.matrix(m$X) %*% t(m$Sol))
  fixedf_variance <- lapply(predictions, function (j) apply(j, MARGIN = 2, var))
  
  # make a dataframe of all the id and untis from all 4 chains
  random_out<-do.call(rbind.data.frame, random)
  fixed_eff <- data.frame(values = as.numeric(unlist(fixedf_variance)))
  
  ## Pull out variance components
  h2 <- random_out$id /
    (random_out$id + random_out$pe + random_out$units + fixed_eff$values)
  
  # permanent environmental effects
  pe2 <- (random_out$pe) /
    (random_out$id + random_out$pe + random_out$units + fixed_eff$values)
  
  # repeatability
  ind2 <- (random_out$id + random_out$pe) /
    (random_out$id + random_out$pe + random_out$units + fixed_eff$values)
  
  #evolvability
  ia <- random_out$id / (mean(na.omit(trait)))^2
  
  h2_df <- data.frame(var = 'h2',
                      median = median(h2),
                      mean = mean(h2),
                      lower = quantile(h2, 0.025),
                      upper = quantile(h2, 0.975),
                      grm = paste(grm.name, name, sep = "_"),
                      stringsAsFactors = FALSE)
  
  pe2_df <- data.frame(var = 'pe2',
                       median = median(pe2),
                       mean = mean(pe2),
                       lower = quantile(pe2, 0.025),
                       upper = quantile(pe2, 0.975),
                       grm = paste(grm.name, name, sep = "_"),
                       stringsAsFactors = FALSE)
  
  ind2_df <- data.frame(var = 'ind2',
                        median = median(ind2),
                        mean = mean(ind2),
                        lower = quantile(ind2, 0.025),
                        upper = quantile(ind2, 0.975),
                        grm = paste(grm.name, name, sep = "_"),
                        stringsAsFactors = FALSE)
  
  ia_df <- data.frame(var = 'ia',
                        median = median(ia),
                        mean = mean(ia),
                        lower = quantile(ia, 0.025, na.rm=TRUE),
                        upper = quantile(ia, 0.975, na.rm=TRUE),
                        grm = paste(grm.name, name, sep = "_"),
                        stringsAsFactors = FALSE)
  
  return(list(h2_df, pe2_df, ind2_df, ia_df))

}


# start time
h2_start <- list()
for (i in 1:length(grm90_chains_start)){
  h2_start[[i]] <- getHeritabilityMig(grm = grm90_chains_start[[i]], 
                                  grm.name = 'grm90_chains_start',
                                  name = i,
                                  trait = df_90[[i]]$start)
}

nr0.05_start <- do.call('rbind', h2_start[[1]])
nr0.1_start <- do.call('rbind', h2_start[[2]])
unf_start <- do.call('rbind', h2_start[[3]])
df_start <- h2_start %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

# end time
h2_end <- list()
for (i in 1:length(grm90_chains_end)){
  h2_end[[i]] <- getHeritabilityMig(grm = grm90_chains_end[[i]], 
                                   grm.name = 'grm90_chains_end',
                                   name = i,
                                   trait = df_90[[i]]$end)
}

nr0.05_end <- do.call('rbind', h2_end[[1]])
nr0.1_end <- do.call('rbind', h2_end[[2]])
unf_end <- do.call('rbind', h2_end[[3]])
df_end <- h2_end %>% 
  lapply(as.data.frame) %>% 
  bind_rows()


# speed (migratory)
h2_speed <- list()
for (i in 1:length(grm90_chains_speed_dist)){
  h2_speed[[i]] <- getHeritabilityMig(grm = grm90_chains_speed_dist[[i]], 
                                      grm.name = 'grm90_chains_speed_dist',
                                      name = i,
                                      trait = df_90[[i]]$log_speed2)
}


nr0.05_speed <- do.call('rbind', h2_speed[[1]])
nr0.1_speed <- do.call('rbind', h2_speed[[2]])
unf_speed <- do.call('rbind', h2_speed[[3]])
df_speed <- h2_speed %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

# march movement rate (non-migratory)
h2_mar_mov <- list()
for (i in 1:length(grm90_chains_mar_mov)){
  h2_mar_mov[[i]] <- getHeritabilityMig(grm = grm90_chains_mar_mov[[i]], 
                                      grm.name = 'grm90_chains_mar_mov',
                                      name = i,
                                      trait = df_90[[i]]$log_mar_speed)
}

df_mar_mov <- h2_mar_mov %>% 
  lapply(as.data.frame) %>% 
  bind_rows()


# duration
h2_duration <- list()
for (i in 1:length(grm90_chains_duration)){
  h2_duration[[i]] <- getHeritabilityMig(grm = grm90_chains_duration[[i]], 
                                   grm.name = 'grm90_chains_duration',
                                   name = i,
                                   trait = df_90[[i]]$duration)
}

nr0.05_duration <- do.call('rbind', h2_duration[[1]])
nr0.1_duration <- do.call('rbind', h2_duration[[2]])
unf_duration <- do.call('rbind', h2_duration[[3]])
df_duration <- h2_duration %>% 
  lapply(as.data.frame) %>% 
  bind_rows()

# distance
h2_dist <- list()
for (i in 1:length(grm90_chains_dist)){
  h2_dist[[i]] <- getHeritabilityMig(grm = grm90_chains_dist[[i]], 
                                      grm.name = 'grm90_chains_dist',
                                      name = i,
                                     trait = df_90[[i]]$log_distance)
}

nr0.05_dist <- do.call('rbind', h2_dist[[1]])
nr0.1_dist <- do.call('rbind', h2_dist[[2]])
unf_dist <- do.call('rbind', h2_dist[[3]])
df_dist <- h2_dist %>% 
  lapply(as.data.frame) %>% 
  bind_rows()


df_mig <- rbind(df_start, df_end, df_dist, df_duration, df_speed, df_mar_mov)
write.csv(df_mig, "output/h2_mig_update2.csv")

```

### Figures

```{r}
##### Figures ----

# read in proportion of variance table
propvar <- read_excel("propvar_table.xlsx", sheet = 1)
propvar_body <- read_excel("propvar_table.xlsx", sheet = 2)

propvar2 <- melt(as.data.table(propvar), id.vars = 'Trait')
propvar2$variable <- factor(propvar2$variable, levels = c('4Fixed', '3VR', '2VPE', '1VA'))
str(propvar2)

propvar_body <- melt(as.data.table(propvar_body), id.vars = 'Trait')
propvar_body$variable <- factor(propvar_body$variable, levels = c('4Fixed', '3VR', '2VPE', '1VA'))
str(propvar_body)

# Bar chart of proportion of variance
traits_body <- c('Chest', 'Leg', 'IFBF', 'Wt')
traits_mov <- c('Start', 'End', 'Dist', 'Dur', 'Mov Rate\n(mig)', 'Mov Rate\n(non-mig)')

p<-ggplot(propvar2, aes(x = Trait, y = value, fill = variable)) +
  geom_col(position = "fill", colour = 'black')+
  xlab(NULL)+
  ylab(NULL)+
  theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    axis.text.y = element_text(size = 16, colour = "black"),
    axis.text.x = element_text(size=16, colour = "black"),
    # legend.title = element_blank(),
    legend.text = element_text(size = 16, colour = "black"),
    axis.title.y = element_text(size = 18),
    legend.title = element_text(size = 18)
  )+
  guides(fill=guide_legend(title="Effect"))+
  scale_fill_manual(values=c("#DDCC77", "#88CCEE","#44AA99", '#117733'),
                    labels = c('Fixed effects', 'Residual', 'Permanent environment', 'Additive genetic'))+
  scale_x_discrete(labels = traits_mov)

p

png('output/propvar.png', height = 6, width = 10, units = "in", res = 400)
print(p)
dev.off()

p_body<-ggplot(propvar_body, aes(x = Trait, y = value, fill = variable)) +
  geom_col(position = "fill", colour = 'black')+
  xlab(NULL)+
  ylab(NULL)+
  theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    axis.text.y = element_text(size = 20, colour = "black"),
    axis.text.x = element_text(size=12, colour = "black"),
    # legend.title = element_blank(),
    legend.text = element_text(size = 18, colour = "black"),
    axis.title.y = element_text(size = 20),
    legend.title = element_text(size = 20,  colour = "black")
  )+
  guides(fill=guide_legend(title="Proportion of Variance"))+
  scale_fill_manual(values=c("#DDCC77", "#88CCEE", '#117733'),
                    labels = c('Fixed Effects', 'Residual', 'Additive genetic'))+
  scale_x_discrete(labels = traits_body)

p_body

png('output/propvar_body.png', height = 6, width = 8.5, units = "in", res = 400)
print(p_body)
dev.off()

# Point and Ci plot of h2 and prop variance
var_ci <- read_csv("output/propvar_mig_update_CI.csv")
var_ci$var <- factor(var_ci$var, levels = c("id", "pe", "units", "fixed"))
var_ci$trait <- factor(var_ci$trait, levels = c("start", "end", "distance", "duration",
                                               "speed", "mar_mov"))

propvar_ci <- ggplot(var_ci, aes(x = var, y = mean, colour = var)) +
  geom_pointrange(
    aes(ymin = lower, ymax = upper),
    linewidth = 0.6
  ) +
  facet_wrap(~ trait, ncol = 1, strip.position = "right",
             labeller = as_labeller(c(
               start = "Start",
               end   = "End",
               distance = "Distance",
               duration = "Duration",
               speed = "Movement rate\n(migration)",
               mar_mov = "Movement rate\n(non-migratory)"
             ))) +
  labs(
    x = "Variance component",
    y = "Proportion of variance (posterior mean with 95% CI)"
  ) +
  theme_bw() +
  scale_x_discrete(
    labels = c(
      id    = "Additive\ngenetic",
      pe    = "Permanent\nenvironment",
      units = "Residual",
      fixed = "Fixed effects"
    ))+
  scale_colour_manual(values=c('#117733', "#44AA99", "#88CCEE","#DDCC77"))+
  theme(
    strip.background = element_blank(),
    strip.text = element_text(face = "bold"),
    legend.position = "none"
  )

png('output/propvar_ci_update.png', height = 8, width = 6, units = "in", res = 400)
print(propvar_ci)
dev.off()


### GRM Figures
filenames <- list.files("input/gcta_grm/popmap4r08_90_indrm", pattern="*.grm.bin", full.names = T)
uf.list <- list()
uf.df <- vector("list", length(filenames))

for(i in 1:length(filenames)){
  # read in the grm list of 4 
  uf.list <- list()
  uf.list <- ReadGRMBin(prefix = substr(filenames[i],1,nchar(filenames[i])-8), AllN = T)
  
  # get name
  name <- substr(filenames[i],44,nchar(filenames[i])-8)
  off <- data.table(uf.list$off)
  diag <- data.table(uf.list$diag)
  N <- max(uf.list$N)
  id <- length(uf.list$id$V2)
  
  moff <- mean(off$V1)
  mdiag <- mean(diag$V1)
  mindiag <- min(diag$V1)
  maxdiag <- max(diag$V1)
  voff <- var(off$V1)
  vdiag <- var(diag$V1)
  
  df <- data.table(name, id, N, moff, voff, mdiag, vdiag, mindiag, maxdiag)
  names(df) <- c('GRM', 'N', 'SNPs', 'mean off', 'var off', "mean diagonal", 'var diagonal', 
                 'min diagonal', 'max diagonal')
  
  uf.df[[i]] <- df
  
}

ufdf <- do.call('rbind', uf.df)
ufdf

grm<-melt(GRM_90[[3]])
setDT(grm)
off<-grm[Var1!=Var2,]
diag<-grm[Var1==Var2,]

p1 <-ggplot(aes(x=value), data = off)+
  geom_histogram(bins=100, colour='#332288', fill = '#332288')+
  ylab(NULL)+
  xlab(NULL)+
  #ggtitle(paste(name, " SNPs = ", N, " N = ", id, ' mean off = ',
  #round(moff, digits = 6), sep = ""))+
  theme(
    axis.text.x = element_text(size=12, colour = "black"),
    axis.text.y = element_text(size=12, colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.grid = element_blank(),
    panel.border = element_rect(colour = 'black', fill = 'transparent'),
    panel.background = element_rect(fill='transparent'), #transparent panel bg
    plot.background = element_rect(fill='transparent', color=NA),
  )

p2 <-ggplot(aes(x=value), data = off)+
  geom_histogram(bins=100, colour='#332288', fill = '#332288')+
  geom_vline(xintercept = 0.1, colour = 'red')+
  ylab(NULL)+
  xlab(NULL)+
  coord_cartesian(ylim=c(0, 50))+
  theme(
    axis.text.x = element_text(size=18, colour = "black"),
    axis.text.y = element_text(size=18, colour = "black"),
    axis.title = element_text(size=20, colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.grid = element_blank(),
    panel.border = element_rect(colour = 'black', fill = 'transparent'),
    panel.background = element_rect(fill='transparent'), #transparent panel bg
    plot.background = element_rect(fill='transparent', color=NA),
  )

p3 <-p2 + annotation_custom(ggplotGrob(p1), xmin = 0.2, xmax = 0.8, 
                            ymin = 20, ymax = 50)
p3

png('output/grm_freq_line.png', height = 3.5, width = 6, units = "in", res = 400, bg='transparent')
print(p3)
dev.off()

```

